No proof nets for \capsabbrev{mll} with units

### 2014-05-16 main/LICS.tex
final version

@@

Later, a striking technical innovation of linear logic was the introduction of proof nets \cite{Girard-1987,Danos-Regnier-1989}, which define a canonical form for proofs in the unitless fragment of multiplicative linear logic---two proofs are equivalent if and only if they have the same proof net---so proof nets offer a simple decision procedure for proofs in this fragment. This naturally raises the question whether proof nets can be extended to work in the presence of units. The work in this direction begins with \cite{Trimble-1994,Blute-Cockett-Seely-Trimble-1996} via \color{red}\cite{Lamarche-Strassburger-2006} \color{black} and perhaps culminates in \cite{HughesFreeStar,HughesMLLProofNets}; but these proof nets are not canonical and must be identified up to a rewiring equivalence.

@@

In this paper we establish that the proof equivalence problem for multiplicative linear logic with units is \capsabbrev{pspace}-complete. It follows that there can be no satisfactory canonical notion of proof net: 
\color{red}
if a notion of proof net were to exist where both the translation from proofs to nets and equality of proof nets are polynomial-time computable, this would imply that \capsabbrev{p}\,=\,\capsabbrev{pspace}.
\color{black}

@@

This is one of a class of problems called \emph{reconfiguration problems} \cite{ReconfigurationProblems}: can one solution to a given problem be transformed into another by a series of elementary changes, while remaining valid throughout?
%
For example, the reconfiguration problem for boolean satisfiability (SAT) asks whether one satisfying assignment can be transformed into another by changing the value of one atomic formula at a time, without passing via a non-satisfying assignment.
%
It is not uncommon for an \capsabbrev{np}-complete problem to have an associated reconfiguration problem that is \capsabbrev{pspace}-complete \cite{ReconfigurationProblems}; one example is SAT-reconfiguration, which is \capsabbrev{pspace}-complete. \capsabbrev{mll} proof equivalence may be regarded as the reconfiguration problem associated with \capsabbrev{mll} proof search, which is \capsabbrev{np}-complete \cite{Kanovich-1992,Lincoln-Winkler-1994}.

\begin{figure}[!b]
\quad\MLLrule b
\hfill\MLLrule 1
\hfill\MLLrule p
\hfill\MLLrule t
\quad
\caption{Inference rules for unit-only \capsabbrev{mll}}
\label{fig:MLL}
\end{figure}

@@

\aftersection
%
The formulae of unit-only multiplicative linear logic are given by the following grammar.
%
\setMidspace{5pt}
\[
	A,B,C \Coloneq \bot \Mid 1 \Mid A\parr B \Mid A\tn B
\]
%
The connectives $\tn$ and $\parr$ will be considered up to associativity, and \emph{duality} $\dual A$ is via DeMorgan.
%
A \emph{sequent} $\Gamma,\Delta$ will be a multiset of formulae.
%
Within a sequent, connectives and units will be \emph{named} with distinct elements from an arbitrary set of names, e.g.\
\[
	\named a1\named b\parr\named c1,\named d\bot\named e\tn\named f\bot~.
\]
%
This allows to 1) identify \emph{occurrences} of subformulae uniquely by the name of their root connective, e.g.\ as $\named bA$, 2) distinguish the two proofs of the above sequent while using standard multiset sequents, and 3) easily extract proof nets, as graphs using the names of connectives as vertices.
%
Names will mostly be left implicit.

@@

The permutations of sequent proofs are exactly the identifications imposed by the categorical semantics of \capsabbrev{mll}, star-autonomous categories \cite{Barr-1991} (and semi-star-autonomous categories \cite{Houston-2008,Heijltjes-Strassburger} for \capsabbrev{mll} without units).
%
Proof equivalence for \capsabbrev{mll} is therefore equivalent to the \emph{word problem} for star-autonomous categories: the problem whether two representations denote the same morphism in any star-autonomous category.

@@

\noindent
An edge $\edge a{\links(a)}$ in a proof net or switching graph is called a \emph{link} or a \emph{jump}.
%
\color{red}
The restriction that jumps must target $\1$-occurrences (rather than any connective) is a convenience---it can be circumvented by replacing a subformula $A$ by the equivalent $A\tn\1$.
\color{black}

\begin{figure}
\color{red}
\[
	\netA 1223 ~\perm*~
	\netA 1323 ~\perm*~
	\netA 1321 ~\perm*~
	\netA 2321
\]
\caption{A rewiring sequence on proof nets}
\label{fig:net rewiring}
\end{figure}

@@

\noindent
\color{red}
An example rewiring sequence is given in Figure~\ref{fig:net rewiring}, using the notation for proof nets introduced below.
\color{black}

@@

\begin{proposition}[\cite{Danos-Regnier-1989}, \color{red}\cite{Lamarche-Strassburger-2006}\color{black}]
\label{prop:correctness and sequentialisation}
%
%
If $\Pi\toNet\links$ and $\Pi$ has conclusion $\Gamma$, then $\links$ is a proof net for $\Gamma$.
%For a proof $\Pi$ with conclusion $\Gamma$, if $\Pi\toNet\links$ then $\links$ is a proof net for $\Gamma$.
%
For a net $\links$ for $\Gamma$, there is a proof $\Pi$ of $\Gamma$ such that $\Pi\toNet\links$ (\emph{sequentialisation}).
%
\end{proposition}

@@

\begin{proposition}[\cite{HughesMLLProofNets}] %[\citeauthor{HughesMLLProofNets}, \citeyear{HughesMLLProofNets}]
\label{prop:proof nets work}
%
For proofs $\Pi$, $\Pi'$ and proof nets $\links$, $\links'$ such that $\Pi\toNet\links$ and $\Pi'\toNet\links'$, $\Pi\perm\Pi'$ if and only if $\links\perm\links'$.
%
\end{proposition}

@@

\aftersubsection
We will use a concise diagrammatic notation for sequents and proof nets.
%
The units $1$ and $\bot$ are represented by a circle~($\circ$) and a disc~($\bullet$) respectively;
%
formulae related by a tensor will be connected by edges; 
formulae related by a par will be juxtaposed, and collected in a box when a par-formula is an immediate subformula of a tensor-formula \color{red}(and also for illustrative purposes)\color{black}.
%
For example, the following denote the same sequent:
\[
	\bot\tn\bot,\bot\tn\bot,\1,(\1\parr\1\parr\1)\tn\bot
\]
\[
\begin{tikzpicture}[x=5mm,y=-5mm,octo]
	\draw (0,0) node[bullet] (a) {} -- (1,0) node[bullet] (b) {} (2,0) node[bullet] (c) {} -- (3,0) node[bullet] (d) {};
	\node[circ] (1) at (4,0) {};
	\node[circ] (2) at (5.25,0) {}; \node[circ] (3) at (6.25,0) {}; \node[circ] (4) at (7.25,0) {};
	\draw[rounded corners] (4.75,-1) rectangle (7.75,1);
	\draw (7.75,0) -- (8.75,0) node[bullet] (e) {};
\end{tikzpicture}
\]
The links of a proof net are added to the sequent as coloured arrows.
%
The following example is a proof net for the above sequent.
\[
\begin{tikzpicture}[x=5mm,y=-5mm,octo]
	\node[circ] (1) at (6,0) {};
	\draw (0,0) node[bullet] (a) {} -- (1,0) node[bullet] (b) {} (3,0) node[bullet] (c) {} -- (4,0) node[bullet] (d) {};
	\node[circ] (2) at (1,3) {}; \node[circ] (3) at (2,3) {}; \node[circ] (4) at (3,3) {};
	\draw[rounded corners] (0,2) rectangle (4,4);
	\draw (4,3) -- (5,3) node[bullet] (e) {};
	\begin{scope}[thepink,->]
			\draw[bend right=10] (a) to (2.center);
			\draw[bend right=10] (b) to (3.center);
			\draw[bend left=10]  (c) to (3.center);
			\draw[bend left=10]  (d) to (4.center);
			\draw[bend right=10] (e) to (1.center);
	\end{scope}
\end{tikzpicture}
\]

We will denote by $A^n$ the sequent consisting of $n$ occurrences of a formula $A$.
%
Given a sequent $\Gamma=A_1,\dotsc,A_n$ we will write $\bigtn\Gamma$ for $A_1\tn\cdots\tn A_n$, and $\bigparr\Gamma$ for $A_1\parr\cdots\parr A_n$.
%
In diagrammatic notation, a big disc will represent an $n$-ary tensor over only $\bot$-formulae, $\bigtn(\bot^n)$, and a big circle will represent an $n$-ary par (or a sequent) over only $1$-formulae, $\bigparr(\1^n)$.
%
A linking between $\bigtn(\bot^n)$ and $\bigparr(\1^n)$ will be represented by a wide arrow, as illustrated below.
\[
\begin{tikzpicture}[x=-5mm,y=5mm,octo]
	\node[big_] (A) at (-1,2) {};
	\node[big1] (!) at (-1,0) {};
	\node at (.5,1) {$\Rightarrow$};
	\draw (2,2) node[bullet] (a) {} -- (3,2) node[bullet] (b) {} -- (4,2) (5,2) -- (6,2) node[bullet] (c) {}; 
	\node[circ] (1) at (2,0) {}; \node[circ] (2) at (3,0) {}; \node[circ] (3) at (6,0) {};
	\draw[dotted] (4,2)--(5,2) (4,0)--(5.1,0);
	\begin{scope}[thepink,->]
			\draw[big>] (A) to (!.center);
			\draw[bend right=10] (a) to (1.center);
			\draw[bend right=10] (b) to (2.center);
			\draw[bend right=10] (c) to (3.center);
	\end{scope}
\end{tikzpicture}
\]
Two formulae $\bigtn(\bot^{i+1})$ and $\bigtn(\bot^{j+1})$ may together connect to a formula $\bigparr(\1^{i+j+1})$, as illustrated below for $i=2$ and $j=3$.
%
Nodes will be labelled $i'$ for $i+1$, so that both the abbreviated formulae and the arithmetic of connecting them remain intuitive.
%
\[
\vc{
  \begin{tikzpicture}[x=5mm,y=-5mm,octo]
	\draw (0,0) node[bullet] (a) {}--
		  (1,0) node[bullet] (b) {}--
		  (2,0) node[bullet] (c) {};
	\draw (4,0) node[bullet] (d) {}--
		  (5,0) node[bullet] (e) {}--
		  (6,0) node[bullet] (f) {}--
		  (7,0) node[bullet] (g) {};
	\foreach \i in {1,...,6} {\node[circ] (\i) at (\i,3) {};}
	\begin{scope}[thepink,->]
			\draw[bend right=10] (a) to (1);
			\draw[bend right=10] (b) to (2);
			\draw[bend right=10] (c) to (3);
			\draw[bend left=10]  (d) to (3);
			\draw[bend left=10]  (e) to (4);
			\draw[bend left=10]  (f) to (5);
			\draw[bend left=10]  (g) to (6);
	\end{scope}
  \end{tikzpicture}
}
\quad\Rightarrow\quad
\vc{
  \begin{tikzpicture}[x=5mm,y=-5mm,octo]
	\node[big_] (A) at (-1,0) {$2'$};
	\node[big_] (B) at ( 1,0) {$3'$};
	\node[big1] (!) at ( 0,3) {$5'$};
	\begin{scope}[thepink,->,big>]
		\draw[bend right=15] (A) to (!);
		\draw[bend left=15]  (B) to (!);
	\end{scope}
  \end{tikzpicture}
}
\]
%
\color{red}
A path in a proof net or switching graph is indicated $\path ab$, and illustrated as below.
\color{black}
%
%A dashed edge will indicate that two vertices are connected, in a proof net or a switching graph, as illustrated below.
\[
\begin{tikzpicture}[x=5mm,y=5mm,octo]
	\node[bullet] (a) at (0,0) {}; \node[bullet] (b) at (2,0) {}; \draw[dashed] (a)--(b);
\end{tikzpicture}
\qquad
\begin{tikzpicture}[x=5mm,y=5mm,octo]
	\node[circ] (a) at (0,0) {}; \node[circ] (b) at (2,0) {}; \draw[dashed] (a)--(b);
	%\coordinate (a') at ($(a) + (.6,0)$); \coordinate (b') at ($(b) + (-.6,0)$);
	%\draw[thepink,dashed] (a')--(b'); \draw[thepink,->] (a')--(a.center); \draw[thepink,->] (b')--(b.center);
\end{tikzpicture}
\]

@@

\aftersection
%
\emph{Non-deterministic constraint logic} \cite{Hearn-Demaine-2005,Demaine-Hearn-2008,GamesPuzzlesAndComputation} is a simple graph-rewriting formalism, used here as a convenient tool for \capsabbrev{pspace}-hardness reduction.
%
A \emph{constraint graph} is a graph with weighted edges and an \emph{inflow constraint}---a natural number that may be taken to be always $2$---on each vertex. A \emph{configuration} is an assignment of directions to the edges of the underlying undirected constraint graph. A rewrite step consists of the reversal of a single edge in a configuration:
rewrite steps must preserve the condition that the total weight of the incoming edges at each vertex is at least its inflow constraint.

\color{red}
Figure~\ref{fig:NCL example} shows an example rewrite sequence in a constraint graph.
%
The central node has inflow constraint 2, the thick blue edge has weight 2, and the thinner red edges have weight 1.
\color{black}

@@

\begin{figure}\color{red}
$
~
\vc{\begin{tikzpicture}[x=-3mm,y=3mm]
	\draw[ncl2,->-] (0,0) -- (3,0);
	\draw[ncl1,->-] (3,0) -- (5,2);
	\draw[ncl1,->-] (3,0) -- (5,-2);
	\node[nclnode] at (3,0) {};
\end{tikzpicture}}
\hfill\perm*\,
\vc{\begin{tikzpicture}[x=-3mm,y=3mm]
	\draw[ncl2,->-] (0,0) -- (3,0);
	\draw[ncl1,-<-] (3,0) -- (5,2);
	\draw[ncl1,->-] (3,0) -- (5,-2);
	\node[nclnode] at (3,0) {};
\end{tikzpicture}}
\hfill\perm*\,
\vc{\begin{tikzpicture}[x=-3mm,y=3mm]
	\draw[ncl2,->-] (0,0) -- (3,0);
	\draw[ncl1,-<-] (3,0) -- (5,2);
	\draw[ncl1,-<-] (3,0) -- (5,-2);
	\node[nclnode] at (3,0) {};
\end{tikzpicture}}
\hfill\perm*\,
\vc{\begin{tikzpicture}[x=-3mm,y=3mm]
	\draw[ncl2,-<-] (0,0) -- (3,0);
	\draw[ncl1,-<-] (3,0) -- (5,2);
	\draw[ncl1,-<-] (3,0) -- (5,-2);
	\node[nclnode] at (3,0) {};
\end{tikzpicture}}
~
$

@@

\begin{figure*}
\color{red}
\[
\vc{\scalebox{0.36}{\BIGexample 3}}
\quad\perm\quad
\vc{\scalebox{0.36}{\BIGexample 2}}
\quad\perm\quad
\vc{\scalebox{0.36}{\BIGexample 1}}
\]
\caption{Rewiring three edge-gadgets connected to a single vertex-gadget}
\label{fig:encoding example}
\end{figure*}

@@

\color{red}
The complete encoding of a constraint graph $G$ will be a sequent $\Gamma$ consisting of:
% 1) all vertex-gadgets combined in a single formula via tensors, 2) all edge-gadgets as individual formulae, 3) a sufficient number of weight absorbers ($\1$-formulae).
\begin{enumerate}
 	\item
all vertex-gadgets, combined in a single formula via tensors, 
	\item
all edge-gadgets as individual formulae, and
	\item
a sufficient number of weight absorbers ($\1$-formulae).
\end{enumerate}
%
A configuration for $G$ will be encoded as a proof net for $\Gamma$, and conversely each proof net for $\Gamma$ may be interpreted as a (partial) configuration for $G$.

@@

\color{black}

@@

\color{red}
The notions and results introduced in this section will be used in the main proofs of the paper, in Section~\ref{sec:correctness}, which show that the encoding of \capsabbrev{ncg}-reconfiguration in \capsabbrev{MLL} proof equivalence, as described informally in the previous section, is correct.
\color{black}

@@

%\noindent
\color{red}
The \emph{ports} of a sub-sequent $\Gamma'$ or subnet $(\Gamma',\links')$ are the root vertices of $\Gamma'$.
%
%The root vertices of $\Gamma'$ are the \emph{ports} of the sub-sequent $\Gamma'$ and of the subnet $(\Gamma',\links')$.
%
For a vertex $v$ naming a par, tensor, or bottom, the subnets of which it is a port correspond to the possible subproofs of the rule introducing $v$ in a sequentialisation (the subproof of a $\1$-subformula must always be empty).
%
%
%A \emph{scope} of a par $\named v\parr$ is a subnet that has $v$ as a port.
%
%The scopes of a par correspond to the possible subproofs of its introduction rule in a sequentialisation of the proof net that it occurs in.

@@

\color{black}

>>>>

\keywords{linear logic, proof equivalence, proof nets, constraint logic, \capsabbrev{pspace}-completeness}

@@

Later, a striking technical innovation of linear logic was the introduction of proof nets \cite{Girard-1987,Danos-Regnier-1989}, which define a canonical form for proofs in the unitless fragment of multiplicative linear logic---two proofs are equivalent if and only if they have the same proof net---so proof nets offer a simple decision procedure for proofs in this fragment. This naturally raises the question whether proof nets can be extended to work in the presence of units. The work in this direction begins with \cite{Trimble-1994,Blute-Cockett-Seely-Trimble-1996} via \cite{Lamarche-Strassburger-2006} and perhaps culminates in \cite{HughesFreeStar,HughesMLLProofNets}; but these proof nets are not canonical and must be identified up to a rewiring equivalence.

@@

In this paper we establish that the proof equivalence problem for multiplicative linear logic with units is \capsabbrev{pspace}-complete. 
%
This effectively rules out the existence of a satisfactory notion of proof net for \capsabbrev{mll} with units---one that reduces proof equivalence to syntactic equality, and where the translation from proofs to proof nets and equality of proof nets are both tractable.

%It follows that there can be no satisfactory canonical notion of proof net: if a notion of proof net were to exist where both the translation from proofs to nets and equality of proof nets are polynomial-time computable, this would imply that \capsabbrev{p}\,=\,\capsabbrev{pspace}.

@@

This is one of a class of problems called \emph{reconfiguration problems} \cite{ReconfigurationProblems}: can one solution to a given problem be transformed into another by a series of elementary changes, while remaining valid throughout?
%
For example, the reconfiguration problem for boolean satisfiability (SAT) asks whether one satisfying assignment can be transformed into another by changing the value of one atomic formula at a time, without passing via a non-satisfying assignment.
%
It is not uncommon for an \capsabbrev{np}-complete problem to have an associated reconfiguration problem that is \capsabbrev{pspace}-complete \cite{ReconfigurationProblems}; an example of this is SAT-reconfiguration. \capsabbrev{mll} proof equivalence may be regarded as the reconfiguration problem associated with \capsabbrev{mll} proof search, which is \capsabbrev{np}-complete \cite{Kanovich-1992,Lincoln-Winkler-1994}.

@@

\aftersection
%
The formulae of unit-only multiplicative linear logic are given by the following grammar.
%
\setMidspace{5pt}
\[
	A,B,C \Coloneq \bot \Mid 1 \Mid A\parr B \Mid A\tn B
\]
%
The connectives $\tn$ and $\parr$ will be considered up to associativity, and \emph{duality} $\dual A$ is via DeMorgan.
%
A \emph{sequent} $\Gamma,\Delta$ will be a multiset of formulae.
%
Within a sequent, connectives and units will be \emph{named} with distinct elements from an arbitrary set of names, e.g.\
\[
	\named a1\named b\parr\named c1,\named d\bot\named e\tn\named f\bot~.
\]
%
This allows to 1) identify \emph{occurrences} of subformulae uniquely by the name of their root connective, e.g.\ as $\named bA$, 2) distinguish the two proofs of the above sequent while using standard multiset sequents, and 3) easily extract proof nets, as graphs using the names of connectives as vertices.
%
Names will often be left implicit.

@@

The permutations of sequent proofs are exactly the identifications imposed by the categorical semantics of \capsabbrev{mll}, star-autonomous categories \cite{Barr-1991} (and semi-star-autonomous categories \cite{Houston-2008,Heijltjes-Strassburger} for \capsabbrev{mll} without units).
%
Proof equivalence for \capsabbrev{mll} is therefore equivalent to the \emph{word problem} for star-autonomous categories: the problem whether two term representations denote the same morphism in any star-autonomous category.

\begin{figure}
\quad\MLLrule b
\hfill\MLLrule 1
\hfill\MLLrule p
\hfill\MLLrule t
\quad
\caption{Inference rules for unit-only \capsabbrev{mll}}
\label{fig:MLL}
\end{figure}

@@

\noindent
An edge $\edge a{\links(a)}$ in a proof net or switching graph is called a \emph{link} or a \emph{jump}.
%
The restriction that jumps must target $\1$-occurrences (rather than any connective) is a convenience---it can be circumvented by replacing a subformula $A$ by the equivalent $A\tn\1$.

\begin{figure}
\[
	\netA 1223 ~\perm*~
	\netA 1323 ~\perm*~
	\netA 1321 ~\perm*~
	\netA 2321
\]
\caption{A rewiring sequence on proof nets}
\label{fig:net rewiring}
\end{figure}

@@

\noindent
An example rewiring sequence is given in Figure~\ref{fig:net rewiring}, using the notation for proof nets introduced below.

@@

\begin{proposition}[\cite{Danos-Regnier-1989}, \cite{Lamarche-Strassburger-2006}]
\label{prop:correctness and sequentialisation}
%
%
If $\Pi\toNet\links$ and $\Pi$ has conclusion $\Gamma$, then $\links$ is a proof net for $\Gamma$.
%For a proof $\Pi$ with conclusion $\Gamma$, if $\Pi\toNet\links$ then $\links$ is a proof net for $\Gamma$.
%
For a net $\links$ for $\Gamma$, there is a proof $\Pi$ of $\Gamma$ such that $\Pi\toNet\links$ (\emph{sequentialisation}).
%
\end{proposition}

@@

\begin{proposition}[\cite{HughesMLLProofNets}] %[\citeauthor{HughesMLLProofNets}, \citeyear{HughesMLLProofNets}]
\label{prop:proof nets work}
%
%If $\Pi\toNet\links$ and $\Pi'\toNet\links'$, then $\Pi\perm\Pi'$ if and only if $\links\perm\links'$.
For proofs $\Pi$, $\Pi'$ and proof nets $\links$, $\links'$, if $\Pi\toNet\links$ and $\Pi'\toNet\links'$, then $\Pi\perm\Pi'$ if and only if $\links\perm\links'$.
%
\end{proposition}

@@

\aftersubsection
We will use a concise diagrammatic notation for sequents and proof nets.
%
The units $1$ and $\bot$ are represented by a circle~($\circ$) and a disc~($\bullet$) respectively;
%
formulae related by a tensor will be connected by edges; 
formulae related by a par will be juxtaposed, and collected in a box when a par-formula is an immediate subformula of a tensor-formula (and also for illustrative purposes).
%
For example, the following denote the same sequent:
\[
	\bot\tn\bot,\bot\tn\bot,\1,(\1\parr\1\parr\1)\tn\bot
\]
\[
\begin{tikzpicture}[x=5mm,y=-5mm,octo]
	\draw (0,0) node[bullet] (a) {} -- (1,0) node[bullet] (b) {} (2,0) node[bullet] (c) {} -- (3,0) node[bullet] (d) {};
	\node[circ] (1) at (4,0) {};
	\node[circ] (2) at (5.25,0) {}; \node[circ] (3) at (6.25,0) {}; \node[circ] (4) at (7.25,0) {};
	\draw[rounded corners] (4.75,-1) rectangle (7.75,1);
	\draw (7.75,0) -- (8.75,0) node[bullet] (e) {};
\end{tikzpicture}
\]
The links of a proof net are added to the sequent as coloured arrows.
%
The following example is a proof net for the above sequent.
\[
\begin{tikzpicture}[x=5mm,y=-5mm,octo]
	\node[circ] (1) at (6,0) {};
	\draw (0,0) node[bullet] (a) {} -- (1,0) node[bullet] (b) {} (3,0) node[bullet] (c) {} -- (4,0) node[bullet] (d) {};
	\node[circ] (2) at (1,3) {}; \node[circ] (3) at (2,3) {}; \node[circ] (4) at (3,3) {};
	\draw[rounded corners] (0,2) rectangle (4,4);
	\draw (4,3) -- (5,3) node[bullet] (e) {};
	\begin{scope}[thepink,->]
			\draw[bend right=10] (a) to (2.center);
			\draw[bend right=10] (b) to (3.center);
			\draw[bend left=10]  (c) to (3.center);
			\draw[bend left=10]  (d) to (4.center);
			\draw[bend right=10] (e) to (1.center);
	\end{scope}
\end{tikzpicture}
\]

We will denote by $A^n$ the sequent consisting of $n$ occurrences of a formula $A$.
%
Given a sequent $\Gamma=A_1,\dotsc,A_n$ we will write $\bigtn\Gamma$ for $A_1\!\tn\cdots\tn A_n$, and $\bigparr\Gamma$ for $A_1\!\parr\cdots\parr\! A_n$.
%
In diagrammatic notation, a big disc will represent an $n$-ary tensor over only $\bot$-formulae, $\bigtn(\bot^n)$, and a big circle will represent an $n$-ary par (or a sequent) over only $1$-formulae, $\bigparr(\1^n)$.
%
A linking between $\bigtn(\bot^n)$ and $\bigparr(\1^n)$ will be represented by a wide arrow, as illustrated below.
\[
\begin{tikzpicture}[x=-5mm,y=5mm,octo]
	\node[big_] (A) at (-1,2) {};
	\node[big1] (!) at (-1,0) {};
	\node at (.5,1) {$\Rightarrow$};
	\draw (2,2) node[bullet] (a) {} -- (3,2) node[bullet] (b) {} -- (4,2) (5,2) -- (6,2) node[bullet] (c) {}; 
	\node[circ] (1) at (2,0) {}; \node[circ] (2) at (3,0) {}; \node[circ] (3) at (6,0) {};
	\draw[dotted] (4,2)--(5,2) (4,0)--(5.1,0);
	\begin{scope}[thepink,->]
			\draw[big>] (A) to (!.center);
			\draw[bend right=10] (a) to (1.center);
			\draw[bend right=10] (b) to (2.center);
			\draw[bend right=10] (c) to (3.center);
	\end{scope}
\end{tikzpicture}
\]
Two formulae $\bigtn(\bot^{i+1})$ and $\bigtn(\bot^{j+1})$ may together connect to a formula $\bigparr(\1^{i+j+1})$, as illustrated below for $i=2$ and $j=3$.
%
Nodes will be labelled $i'$ for $i+1$, so that both the abbreviated formulae and the arithmetic of connecting them remain intuitive.
%
\[
\vc{
  \begin{tikzpicture}[x=5mm,y=-5mm,octo]
	\draw (0,0) node[bullet] (a) {}--
		  (1,0) node[bullet] (b) {}--
		  (2,0) node[bullet] (c) {};
	\draw (4,0) node[bullet] (d) {}--
		  (5,0) node[bullet] (e) {}--
		  (6,0) node[bullet] (f) {}--
		  (7,0) node[bullet] (g) {};
	\foreach \i in {1,...,6} {\node[circ] (\i) at (\i,3) {};}
	\begin{scope}[thepink,->]
			\draw[bend right=10] (a) to (1);
			\draw[bend right=10] (b) to (2);
			\draw[bend right=10] (c) to (3);
			\draw[bend left=10]  (d) to (3);
			\draw[bend left=10]  (e) to (4);
			\draw[bend left=10]  (f) to (5);
			\draw[bend left=10]  (g) to (6);
	\end{scope}
  \end{tikzpicture}
}
\quad\Rightarrow\quad
\vc{
  \begin{tikzpicture}[x=5mm,y=-5mm,octo]
	\node[big_] (A) at (-1,0) {$2'$};
	\node[big_] (B) at ( 1,0) {$3'$};
	\node[big1] (!) at ( 0,3) {$5'$};
	\begin{scope}[thepink,->,big>]
		\draw[bend right=15] (A) to (!);
		\draw[bend left=15]  (B) to (!);
	\end{scope}
  \end{tikzpicture}
}
\]
%
A path in a proof net or switching graph is indicated $\path ab$, and illustrated as below.
%
\[
\begin{tikzpicture}[x=5mm,y=5mm,octo]
	\node[bullet] (a) at (0,0) {}; \node[bullet] (b) at (2,0) {}; \draw[dashed] (a)--(b);
\end{tikzpicture}
\qquad
\begin{tikzpicture}[x=5mm,y=5mm,octo]
	\node[circ] (a) at (0,0) {}; \node[circ] (b) at (2,0) {}; \draw[dashed] (a)--(b);
\end{tikzpicture}
\]

@@

\aftersection
%
\emph{Non-deterministic constraint logic} \cite{Hearn-Demaine-2005,Demaine-Hearn-2008,GamesPuzzlesAndComputation} is a simple graph-rewriting formalism, used here as a convenient tool for \capsabbrev{pspace}-hardness reduction.
%
A \emph{constraint graph} is a graph with weighted edges and an \emph{inflow constraint}---a natural number that may be taken to be always $2$---on each vertex. A \emph{configuration} is an assignment of directions to the edges of the underlying undirected constraint graph. A rewrite step consists of the reversal of a single edge in a configuration, while preserving the condition that the total weight of the incoming edges at each vertex is at least its inflow constraint.

Figure~\ref{fig:NCL example} shows an example rewrite sequence in a part of a constraint graph.
%
The central node has inflow constraint 2, the thick blue edge has weight 2, and the thinner red edges have weight 1.

@@

\begin{figure}
$
~
\vc{\begin{tikzpicture}[x=-3mm,y=3mm]
	\draw[ncl2,->-] (0,0) -- (3,0);
	\draw[ncl1,->-] (3,0) -- (5,2);
	\draw[ncl1,->-] (3,0) -- (5,-2);
	\node[nclnode] at (3,0) {};
\end{tikzpicture}}
\hfill\perm*\,
\vc{\begin{tikzpicture}[x=-3mm,y=3mm]
	\draw[ncl2,->-] (0,0) -- (3,0);
	\draw[ncl1,-<-] (3,0) -- (5,2);
	\draw[ncl1,->-] (3,0) -- (5,-2);
	\node[nclnode] at (3,0) {};
\end{tikzpicture}}
\hfill\perm*\,
\vc{\begin{tikzpicture}[x=-3mm,y=3mm]
	\draw[ncl2,->-] (0,0) -- (3,0);
	\draw[ncl1,-<-] (3,0) -- (5,2);
	\draw[ncl1,-<-] (3,0) -- (5,-2);
	\node[nclnode] at (3,0) {};
\end{tikzpicture}}
\hfill\perm*\,
\vc{\begin{tikzpicture}[x=-3mm,y=3mm]
	\draw[ncl2,-<-] (0,0) -- (3,0);
	\draw[ncl1,-<-] (3,0) -- (5,2);
	\draw[ncl1,-<-] (3,0) -- (5,-2);
	\node[nclnode] at (3,0) {};
\end{tikzpicture}}
~
$

@@

\begin{figure*}
\[
\vc{\scalebox{0.36}{\BIGexample 3}}
\quad\perm\quad
\vc{\scalebox{0.36}{\BIGexample 2}}
\quad\perm\quad
\vc{\scalebox{0.36}{\BIGexample 1}}
\]
\caption{Rewiring three edge-gadgets connected to a single vertex-gadget}
\label{fig:encoding example}
\end{figure*}

@@

The complete encoding of a constraint graph $G$ will then be a sequent $\Gamma$ consisting of:
% 1) all vertex-gadgets combined in a single formula via tensors, 2) all edge-gadgets as individual formulae, 3) a sufficient number of weight absorbers ($\1$-formulae).
\begin{enumerate}
 	\item
all vertex-gadgets, combined in a single formula via tensors, 
	\item
all edge-gadgets as individual formulae, and
	\item
a sufficient number of weight absorbers ($\1$-formulae).
\end{enumerate}
%
A configuration for $G$ will be encoded as a proof net for $\Gamma$, and conversely each proof net for $\Gamma$ may be interpreted as a (partial) configuration for $G$.

@@

@@

The notions and results introduced in this section will be used in the main proofs of the paper, in Section~\ref{sec:correctness}, which show that the encoding of \capsabbrev{ncg}-reconfiguration in \capsabbrev{MLL} proof equivalence, as described informally in the previous section, is correct.

@@

The \emph{ports} of a sub-sequent $\Gamma'$ or subnet $(\Gamma',\links')$ are the root vertices of $\Gamma'$.
%
For a vertex $v$ naming a par, tensor, or bottom, the subnets of which it is a port correspond to the possible subproofs of the rule introducing $v$ in a sequentialisation (the subproof of a $\1$-subformula must always be empty).

@@

In a proof net, the \emph{kingdom} and the \emph{empire} of a vertex $v$ are respectively the smallest and largest subnet that have $v$ as a port.
%
In working with the rewiring relation, the notion of empire can be particularly useful.
%
We will denote the empire of $v$ in $\links$ by $\emp v$ (used both as a graph and a set of vertices).

\begin{proposition}[{\cite[Proposition 2.b]{Bellin-vandeWiele-1995}}]
\label{prop:empire propagation}
The empire $\emp v$ is determined by propagation from $v$: 1) through links; 2) up towards subformulae; 3) into a tensor if one of its subformulae is in $\emp v-\{v\}$; 4) into a par if all its subformulae are in $\emp v-\{v\}$.
\end{proposition}

\noindent
The following three lemmata show how empires are connected to rewiring.
%
Firstly, a jump from $\named v\bot$ may be rewired to exactly those $\1$-occurences that are in the empire of $v$ (Lemma~\ref{lem:rewiring within empire}).
%
Secondly, rewiring this jump preserves the empire of $v$, up to that rewiring (Lemma~\ref{lem:rewiring preserves empire}).
%
Thirdly, the empire of any other vertex $w$ will grow or shrink, or neither, but not both (Lemma~\ref{lem:rewiring affects empires}).
